---
title: "Stats 67 Winter HW 7"
format: html
---

# Homework 7

## Cheating with AI

2-proportion confidence interval steps
```{r}
x2019 <- 96
n2019 <- 96 + 1223

x2023 <- 38
n2023 <- 38 + 212

p2019 <- x2019 / n2019
p2023 <- x2023 / n2023

# Point estimate
p2019 - p2023

# Critical value
(crit_val90 <- qnorm((1 - 0.9) / 2, lower.tail = FALSE))

# Confidence interval standard error
(ci_se <- sqrt(p2019 * (1 - p2019) / n2019 + p2023 * (1 - p2023) / n2023))

# 90% confidence interval

round(
  p2019 - p2023 + c(-1, 1) * crit_val90 * ci_se,
  4
)

# reversed 90% CI
round(
  p2023 - p2019 + c(-1, 1) * crit_val90 * ci_se,
  4
)
```

2-proportion hypothesis test
```{r}
# pooled proportion
p_pooled <- (x2019 + x2023) / (n2019 + n2023)
round(p_pooled, 4)

# pooled std error
pooled_se <- sqrt(p_pooled * (1 - p_pooled) * (1 / n2019 + 1 / n2023))
round(pooled_se, 4)

# One tail of p-value
signif(pnorm(p2019 - p2023, mean = 0, sd = pooled_se), digits = 4)

# p-value calculation
signif(pnorm(p2019 - p2023, mean = 0, sd = pooled_se) * 2, digits = 4)

```

1-mean hypothesis test
```{r}
mu <- 3
n <- 100
x_bar <- 2.75
s <- 0.25

# standard error of x_bar
(x_bar_se <- s / sqrt(n))

# One tail of p-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1), digits = 4)

# P-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1) * 2, digits = 4)

## Change s
s <- 2.5

# standard error of x_bar
(x_bar_se <- s / sqrt(n))

# P-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1) * 2, digits = 4)
```

## Sleep deprivation
2-proportion confidence interval steps
```{r}
x_cal <- 924
n_cal <- x_cal + 10621

x_wash <- 2188
n_wash <- x_wash + 19908

p_cal <- x_cal / n_cal
p_wash <- x_wash / n_wash

# Point estimate
round(p_cal - p_wash, 4)

# Critical value
crit_val90 <- qnorm((1 - 0.9) / 2, lower.tail = FALSE)
round(crit_val90, 4)

# Confidence interval standard error
ci_se <- sqrt(p_cal * (1 - p_cal) / n_cal + p_wash * (1 - p_wash) / n_wash)
round(ci_se, 4)

# 90% confidence interval

round(
  p_cal - p_wash + c(-1, 1) * crit_val90 * ci_se,
  4
)

# reversed 90% CI
round(
  p_wash - p_cal + c(-1, 1) * crit_val90 * ci_se,
  4
)
```

2-proportion hypothesis test
```{r}
# pooled proportion
p_pooled <- (x_cal + x_wash) / (n_cal + n_wash)
round(p_pooled, 4)

# pooled std error
pooled_se <- sqrt(p_pooled * (1 - p_pooled) * (1 / n_cal + 1 / n_wash))
round(pooled_se, 4)

# One tail of p-value
signif(pnorm(p_cal - p_wash, mean = 0, sd = pooled_se), digits = 4)

# p-value calculation
signif(pnorm(p_cal - p_wash, mean = 0, sd = pooled_se) * 2, digits = 4)

```

1-mean hypothesis test
```{r}
mu <- 8
n <- 100
x_bar <- 7.75
s <- 0.25

# standard error of x_bar
(x_bar_se <- s / sqrt(n))

# One tail of p-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1), digits = 4)

# P-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1) * 2, digits = 4)

## Change s
s <- 2.5

```


## Electric vehicle adoption

2-proportion confidence interval steps
```{r}
x_grad <- 8
n_grad <- 29 + x_grad

x_nograd <- 25
n_nograd <- 51 + x_nograd

p_grad <- x_grad / n_grad
p_nograd <- x_nograd / n_nograd

# Point estimate
p_grad - p_nograd

# Critical value
(crit_val90 <- qnorm((1 - 0.9) / 2, lower.tail = FALSE))

# Confidence interval standard error
ci_se <- sqrt(p_grad * (1 - p_grad) / n_grad + p_nograd * (1 - p_nograd) / n_nograd)
round(ci_se, 4)

# 90% confidence interval

round(
  p_grad - p_nograd + c(-1, 1) * crit_val90 * ci_se,
  4
)

# reversed 90% CI
round(
  p_nograd - p_grad + c(-1, 1) * crit_val90 * ci_se,
  4
)
```

2-proportion hypothesis test
```{r}
# pooled proportion
p_pooled <- (x_grad + x_nograd) / (n_grad + n_nograd)
round(p_pooled, 4)

# pooled std error
pooled_se <- sqrt(p_pooled * (1 - p_pooled) * (1 / n_grad + 1 / n_nograd))
round(pooled_se, 4)

# One tail of p-value
signif(pnorm(p_grad - p_nograd, mean = 0, sd = pooled_se), digits = 4)

# p-value calculation
signif(pnorm(p_grad - p_nograd, mean = 0, sd = pooled_se) * 2, digits = 4)

```

1-mean hypothesis test
```{r}
mu <- 300
n <- 100
x_bar <- 275
s <- 25

# standard error of x_bar
(x_bar_se <- s / sqrt(n))

# One tail of p-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1), digits = 4)

# P-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1) * 2, digits = 4)

## Change s
s <- s * 10

# standard error of x_bar
(x_bar_se <- s / sqrt(n))

# P-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1) * 2, digits = 4)
```



# standard error of x_bar
(x_bar_se <- s / sqrt(n))

# P-value
signif(pt((x_bar - mu) / x_bar_se, df = n - 1) * 2, digits = 4)
