---
title: "Data wrangling"
format: html
editor: source
---

```{r}
#| message: false
library(here)
library(tidyverse)
```

```{r}
hw_setup <- read_csv(
  here("data", "hw-setup.csv"),
  show_col_types = FALSE
)
```

## For quantitiative analysis of homework grades

```{r}
# To fix hw 1 section b late penalty
lecb_hw1_q1_11_fix <- data.frame(
  unfixed_total_score = c(6.87, 7.225),
  penalty = c(.5, .5),
  fixed_question_1_11_score = c(2.49, 1.49)
)
```


We want students' total grades and sections for each homework assignment as one csv.

Each assignment has some questions dropped from the total score, the reason for each is noted in hw-setup.csv.

```{r}
all_hw_grades <- NULL

for(i in 1:nrow(hw_setup)) {
  hw_i_og <- read_csv(
    here(
      "data", "deidentified-student-data", "deidentified-files", 
      hw_setup$file_name[i]
    ),
    show_col_types = FALSE
  )
  
  if (!is.na(hw_setup$questions_to_drop[i])) {
    hw_i <- hw_i_og |> 
      select(!contains(str_split_1(
        hw_setup$questions_to_drop[i], 
        pattern = ", "
      )))
    
  } else if (hw_setup$hw[i] == 1 & hw_setup$section[i] == "B") { 
    # Fix the two late penalties
    hw_i <- hw_i_og |> 
      mutate(question_1_11_score = case_when(
        total_score == lecb_hw1_q1_11_fix[1, "unfixed_total_score"] ~ lecb_hw1_q1_11_fix[1, "fixed_question_1_11_score"],
        total_score == lecb_hw1_q1_11_fix[2, "unfixed_total_score"] ~ lecb_hw1_q1_11_fix[2, "fixed_question_1_11_score"],
        TRUE ~ question_1_11_score
      )) |> 
      mutate(total_score = case_when(
        total_score == lecb_hw1_q1_11_fix[1, "unfixed_total_score"] ~ total_score / lecb_hw1_q1_11_fix[1, "penalty"],
        total_score == lecb_hw1_q1_11_fix[2, "unfixed_total_score"] ~ total_score / lecb_hw1_q1_11_fix[2, "penalty"],
        TRUE ~ total_score
      ))
  }
  
  # Need to sum across the score columns to compute hw grade
  score_cols <- colnames(hw_i)[
    str_detect(colnames(hw_i), pattern = "score") & 
    str_detect(colnames(hw_i), pattern = "total", negate = TRUE)
  ]
  
  hw_i$total_score_adj <- rowSums(hw_i[, score_cols], na.rm = TRUE)
  
  # Need to sum across the weight columns to compute max points possible
  max_points_cols <- colnames(hw_i)[
    str_detect(colnames(hw_i), pattern = "weight")
  ]
  
  hw_i$max_points_adj <- rowSums(hw_i[max_points_cols], na.rm = TRUE)
  
  hw_i_cleaned <- hw_i |> 
    filter(status != "Missing") |> 
    select(
      random_id, 
      total_score, 
      total_score_adj, 
      max_points, 
      max_points_adj, 
      status
    ) |>
    mutate(section = hw_setup$section[i]) |> 
    mutate(hw = hw_setup$hw[i]) |> 
    mutate(context = hw_setup$context[i]) |> 
    mutate(choice = hw_setup$choice[i])
    
  
  all_hw_grades <- all_hw_grades |> 
    bind_rows(hw_i_cleaned)
}

all_hw_grades <- all_hw_grades |> 
  rename(total_score_pts_adj = total_score_adj) |> 
  mutate(total_score_per = 100 * total_score_pts_adj / max_points_adj) |> 
  arrange(section, random_id, hw)

# Add average homework grade prior to intervention (Homeworks 1-4)
all_hw_grades <- all_hw_grades |> 
  filter(hw %in% c(1:4)) |> 
  group_by(random_id) |> 
  summarize(ave_hw_per_pre_intervention = mean(total_score_per, na.rm = TRUE)) |> 
  full_join(all_hw_grades, by = "random_id") |> 
  mutate(hw_total_score_diff = total_score_per - ave_hw_per_pre_intervention)
```

### Clean individual issues

#### Two assignments for one student
One student has two assignments for week 9.
The 2003 movies submission is the fully attempted submission, while the bike shares submission was barely attempted, so we drop that one.
```{r}
duplicated_id <- all_hw_grades |> 
  filter(hw == 9, context %in% c("Bike shares", "2003 movies")) |> 
  filter(duplicated(random_id)) |> 
  pull(random_id)

all_hw_grades <- all_hw_grades %>%
  filter(!(random_id == duplicated_id & hw == 9 & context == "Bike shares"))
```

### Save cleaned data
```{r}
all_hw_grades |> 
  arrange(section, random_id, hw) |> 
  write_csv(here("data", "cleaned-data", "hw-grades.csv"))
```

## For qualitiave analysis about data context

We want one csv for students with choice of their responses about why they chose the data context they chose

```{r}
all_context_reasons <- NULL
intervention_hws <- which(hw_setup$hw %in% 6:9)
hws_not_read_in <- NULL

for(i in intervention_hws) {
  context_reason_i_og <- read_csv(
    here(
      "data", "deidentified-student-data", "deidentified-files", 
      hw_setup$file_name[i]
    ),
    show_col_types = FALSE
  )
  
  no_data_context_qs <- is.na(hw_setup$data_context_questions[i])
  hw_unchosen <- sum(context_reason_i_og$status == "Graded") == 0
  
  if (no_data_context_qs | hw_unchosen) {
    hws_not_read_in <- c(hws_not_read_in, hw_setup$file_name[i])
    
  } else {
    context_reason_i <- context_reason_i_og |> 
      rename(data_context_question = paste0(
        hw_setup$data_context_questions[i], "_response"
      )) |> 
      filter(status != "Missing") |> 
      select(random_id, data_context_question) |> 
      mutate(section = hw_setup$section[i]) |> 
      mutate(hw = hw_setup$hw[i]) |> 
      mutate(context = hw_setup$context[i]) |> 
      mutate(choice = hw_setup$choice[i])
    
    if (hw_setup$choice[i]) {
    context_reason_i <- context_reason_i |> 
      mutate(q_end_removed = str_split_i(
        data_context_question, 
        pattern =  '\\"\\}', 
        1
      )) |> 
      mutate(q_start_removed = str_split_i(
        q_end_removed, 
        pattern =  "\\{\"0\"=>\\\"", 
        2
      )) |> 
      mutate(context_q1_response = str_split_i(
        q_start_removed, 
        pattern =  "\", \\\"1\\\"=>\\\"", 
        1
      )) |> 
      mutate(context_q2_response = str_split_i(
        q_start_removed, 
        pattern =  "\", \\\"1\\\"=>\\\"", 
        2
      )) |> 
      mutate(context_q3_response = "not applicable") |> 
      select(!c(q_end_removed, q_start_removed))
    
    } else {
      context_reason_i <- context_reason_i |> 
        mutate(context_q1_response = "not applicable") |> 
        mutate(context_q2_response = "not applicable") |> 
        mutate(context_q3_response = data_context_question)
    }
  
    all_context_reasons <- all_context_reasons |> 
      bind_rows(context_reason_i)
  
  }
  
}

all_context_reasons <- all_context_reasons |> 
  mutate(assignment_id = str_c(hw, section, random_id)) |> 
  select(
    assignment_id, 
    random_id, 
    hw, 
    section, 
    choice, 
    context,
    context_q1_response, 
    context_q2_response, 
    context_q3_response
  ) |> 
  mutate(ds_context1_theme = NA) |> 
  mutate(ds_context1_theme_reason = NA) |> 
  mutate(ds_context2_theme = NA) |> 
  mutate(ds_context2_theme_reason = NA) |> 
  mutate(ds_context3_theme = NA) |> 
  mutate(ds_context3_theme_reason = NA) |> 
  mutate(nonspecific_theme = NA) |> 
  mutate(nonspecific_theme_reason = NA) |> 
  mutate(nontheme_notes = NA)
```

One student has two assignments for week 9.
The 2003 movies submission is the fully attempted submission, while the bike shares submission was barely attempted, so we drop that one.
```{r}
duplicated_id <- all_context_reasons |> 
  filter(hw == 9, context %in% c("Bike shares", "2003 movies")) |> 
  filter(duplicated(random_id)) |> 
  pull(random_id)

all_context_reasons <- all_context_reasons %>%
  filter(!(random_id == duplicated_id & hw == 9 & context == "Bike shares"))
```

### Save cleaned data
```{r}
all_context_reasons |> 
  arrange(section, hw, context) |> 
  write_csv(here("data", "cleaned-data", "hw-context-reasons-themes-Catalina.csv"))

all_context_reasons |> 
  arrange(section, hw, context) |> 
  write_csv(here("data", "cleaned-data", "hw-context-reasons-themes-Mine.csv"))
```

We want another csv for students without choice of their responses about what the liked and didn't like about data context

## For qualitative analysis of pre and post-study surveys

We want one csv for responses to post-study survey

```{r}
post_survey_lec_a <- read_csv(
  here("data", "01-raw-deidentified-student-data", "leca-post-survey-deidentified.csv"),
  show_col_type = FALSE
) |> 
  mutate(section = "A")

post_survey_lec_b <- read_csv(
  here("data", "01-raw-deidentified-student-data", "lecb-post-survey-deidentified.csv"),
  show_col_type = FALSE
) |> 
  mutate(section = "B")

post_survey <- post_survey_lec_a |> 
  bind_rows(post_survey_lec_b) |> 
  filter(status == "Graded") 

choice_opinions <- post_survey |> 
  select(section, random_id, student_response = question_10_response) |> 
  mutate(
    negative_theme = NA,
    negative_theme_reason = NA,
    impartial_theme = NA,
    impartial_theme_reason = NA,
    positive_theme = NA,
    positive_theme_reason = NA,
    nonspecific_theme = NA,
    nonspecific_theme_reason = NA
  )
```

Save a file for each reviewer
```{r}
write_csv(
  choice_opinions, 
  here("data", "02-independent-reviews", "survey-choice-themes-Cat.csv")
)

write_csv(
  choice_opinions, 
  here("data", "02-independent-reviews", "survey-choice-themes-Mine.csv")
)
```

